Critical Transitions in the Spatial Synchrony of Metapopulation Dynamics

### 2020-09-17 writing/manuscript.tex
updating
updates sept17

\title{Critical Transitions in Metapopulation Connectivity}
\author[1,2]{M.D. Catchen}
\author[1]{S.M. Flaxman}
\affil[1]{\small{Department of Ecology and Evolutionary Biology, University of Colorado at Boulder}}
\affil[2]{\small{Department of Biology, McGill University}}
\date{\today}

@@

Landscape connectivity intro

Ecological processes are inherently the product of interactions across all scales of biological organization \cite{levin_problem_1992}. This process of \textit{emergence}, by which parts come together to form a whole with properties that don't exist among the individual parts, has been studied across a wide variety of disciplines \cite{manrubia_emergence_2004} and is a ubiquitous phenomenon in complex systems. One potential cause of emergent behavior is \textit{synchrony} among the individual parts of a complex system. When many independent parts come together to act as a whole, their dynamics become synchronized.
This behavior is ubiquitous in biological systems across all scale of organization.
From collections of cells acting together: the heart beating in rhythm \cite{womelsdorf_modulation_2007}, neurons firing in unison \cite{strogatz_sync_2003}---to behavior among organisms: the flash of fireflies \cite{otte_theories_1980} or the migration of birds \cite{spottiswoode_extrapair_2004}---to interactions between organisms: synchrony between abundances of predators and prey \cite{lv}, and of phenology \cite{van_asch_phenology_2007,burkle_future_2011}.
Synchrony, by definition, involves different entities changing over time in the same way. Within ecology, there has long been a focus on spatial synchrony, that is—how does spatial distribution of ecological entities affect whether they change together or separately? \cite{jarillo_spatial_2020, kendall_dispersal_2000, hanski_spatial_1993}.
This is, in large part, due to the applied importance of understanding the effect of habitat loss on natural populations. Many theoretical studies have shown the two primary factors that develop spatial synchrony across space are dispersal and environmental covariance \cite{ripa_analysing_2000, abbott_does_2007}.
Within this theory, one maxim that has developed is \textit{Moran's rule}, which states that spatial synchrony is proportional to the covariance in the environmental conditions across space \cite{ranta_synchrony_1995, bjornstad_spatial_1999}.

@@

>>>>

\title{Critical Transitions in the Spatial Synchrony of Metapopulation Dynamics}
\author[1,2]{M.D. Catchen}
\author[1]{S.M. Flaxman}
\affil[1]{\small{Department of Ecology and Evolutionary Biology, University of Colorado at Boulder}}
\affil[2]{\small{Department of Biology, McGill University}}
\date{\today}

@@

Anthropogenic activity has rapidly reshaped Earth's habitats.

To mitigate the known contribution of habitat loss to declining biodiversity \cite{}, much attention has been paid to promoting \textit{landscape connectivity}.

It is still largely unknown how ecological systems respond to habitat loss and enduing subdivision.

The metapopulation framework has sought to answer this question.

Spatial synchrony

What is the spatial scale that constitutes a `single' population?
How does landscape connectivity effect the scale of spatial synchrony?

Review of synchrony and literature.
  Spatial Synchrony
  Moran's Rule
  Dispersal Induced Stability

Better understand landscape structure and synchrony, as a step toward understanding
planning landscapes to minimize the probability of extinction.

Ecological processes are inherently the product of interactions across all scales of biological organization \cite{levin_problem_1992}. This process of \textit{emergence}, by which parts come together to form a whole with properties that don't exist among the individual parts, has been studied across a wide variety of disciplines \cite{manrubia_emergence_2004} and is a ubiquitous phenomenon in complex systems. One potential cause of emergent behavior is \textit{synchrony} among the individual parts of a complex system. When many independent parts come together to act as a whole, their dynamics become synchronized.
This behavior is ubiquitous in biological systems across all scale of organization.
From collections of cells acting together: the heart beating in rhythm \cite{womelsdorf_modulation_2007}, neurons firing in unison \cite{strogatz_sync_2003}---to behavior among organisms: the flash of fireflies \cite{otte_theories_1980} or the migration of birds \cite{spottiswoode_extrapair_2004}---to interactions between organisms: synchrony between abundances of predators and prey \cite{lv}, and of phenology \cite{van_asch_phenology_2007,burkle_future_2011}.

Synchrony, by definition, involves different entities changing over time in the same way. Within ecology, there has long been a focus on spatial synchrony, that is—how does spatial distribution of ecological entities affect whether they change together or separately? \cite{jarillo_spatial_2020, kendall_dispersal_2000, hanski_spatial_1993}.
This is, in large part, due to the applied importance of understanding the effect of habitat loss on natural populations. Many theoretical studies have shown the two primary factors that develop spatial synchrony across space are dispersal and environmental covariance \cite{ripa_analysing_2000, abbott_does_2007}.
Within this theory, one maxim that has developed is \textit{Moran's rule}, which states that spatial synchrony is proportional to the covariance in the environmental conditions across space \cite{ranta_synchrony_1995, bjornstad_spatial_1999}.

@@

Does eigenvector centrality predict syncrhony?
